Data Specialization for a Staged Lambda Calculus

### 2014-11-11 paper/splitting.tex
various fixes
added residual++ splitting subection

Recall that partial values and residual tables produced by \bbone-evaluation may
contain content at \emph{both} stages. For example the term:
\begin{lstlisting}
1`(1+2, next{`2`hold{`1`3+4`2`} + 5`1`})`
\end{lstlisting}
evaluates via $\redonesym$ to:
\begin{lstlisting}
2`[yhat|->7+5]` 1`(3,next{`2`yhat`1`})`
\end{lstlisting}
To match the results of \bbone-evaluation,
splitting the term should generate an~$e_1$ that reduces to the stage~\bbone\ content of the partial value, \texttt{3},
and an~$e_2$ corresponding to the stage~\bbtwo\ computation: \\
$\letin \yhat {\mathtt{7+5}} \yhat$.

@@

>>>>

Recall that partial values and residual tables produced by \bbone-evaluation may
contain content at \emph{both} stages. For example the term:
\begin{lstlisting}
1`(1+2, next{`2`7 + 5`1`})`
\end{lstlisting}
evaluates via $\redonesym$ to:
\begin{lstlisting}
2`[yhat|->7+5]` 1`(3,next{`2`yhat`1`})`
\end{lstlisting}
To match the results of \bbone-evaluation,
splitting the term should generate an~$e_1$ that reduces to the stage~\bbone\ content of the partial value, \texttt{3},
and an~$e_2$ corresponding to the stage~\bbtwo\ computation: \\
$\letin \yhat {\mathtt{7+5}} \yhat$.

@@

\subsection{Improved Definition}
\label{sec:resumers}

The definitions of masking and splitting can be augmented to better handle special cases.
One such special case is the behavior of stage \bbone\ splitting at stage \bbone\ types that contain no $\fut$s or $\to$s.
We call such types {\em monostage} (judgmentally, $A~mono$), because a partial value $v$ with a monostage type can contain no $\next$s.

Intuitively, for $[\xi;v] : A$ and $A~mono$, we would expect that $\maskt{\xi;v}$ only reduce to something ``trivial."
There are two reasonable definitions of trivial here: 
\begin{enumerate}
\item a value is trivial if it is exactly \texttt{()},
\item a value is trivial if it is in the closure of \texttt{()} and \texttt{(-,-)}
\end{enumerate}
Indeed, as $\maskt{-;-}$ is currently defined, this property holds for the latter definition of trivial,
but we can make it hold for the former definition by adjusting just the tuple rule:
\[
\maskt{\cdot;(v_1,v_2)} = \left\{ \begin{array}{ll} 
\mathtt{()} &\text{if } (v_1,v_2):A~mono\\ 
(\maskt{\cdot;v_1},\maskt{\cdot;v_2}) &\text{otherwise} \end{array} \right.
\]
From here, the splitting rules can be made to comport with this new 
definition adding special-case tuple, projection, and application 
rules\footnote{The old tuple, projection, and application rules 
must also be augmented to apply only when the $mono$ premises don't hold.
Also, all splitting rules need to be augmented to carry types, though
we omit the contexts here.}:
\begin{mathpar}
\infer
	{(e_1,e_2):A\times B\splitonesym [\ldots,(l_1,l_2).(r_1;r_2;())]}
	{e_1 : A \splitonesym [c_1,l_1.r_1] 
	& e_2 : B \splitonesym [c_2,l_2.r_2] 
	& A \times B~mono} 
\and
\infer
	{\pio~e:A\splitonesym [\ldots,l.(r;())]}
	{e : A\times B \splitonesym [c,l.r] 
	& A \times B~mono} 
\and
\infer
	{e_1~e_2: B\splitonesym [\ldots,(l_1,l_2,l_3).(r_1 (r_2,l_3); ())]}
	{e_1 : A \to B \splitonesym [c_1,l_1.r_1] 
	& e_2 : A \splitonesym [c_2,l_2.r_2] 
	& B~mono} 
\end{mathpar}
